---
title: "Covid vaccine regression"
author: ""
date: "03/08/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

## Regression problem

- We will run regression and other related models for Covid-19 vaccination data

## Libiraries

- We will use the following packages

```{r}
library(tidyverse)
library(caret)
library(glmnet)
```

## Load data

We will use the following data. It is a combined dataset from three data sources we have been using. The code for processing is available at `data_prep/data_preparation.R`.

```{r}
data_vac <- read_csv("data/vaccine-data.csv.gz") 
```


## Check data

Let's have a cursory look at the data, especially check the distribution of the output variable `Booster_Doses_18Plus_Vax_Pct` Do we need conversion?

### `head()`

```{r}

```

### Check the distribution of the output

```{r}

```

```{r}

```


## Decide the variable to include as input

- There are 47 variables what are possible predictors? Especially:
  - trump_pct
  - demography: TotalPop, Men, Women, Hispanic, White, Black, Native, Asian, Pacific, VotingAgeCitizen, Income, IncomePerCap, Poverty, ChildPoverty, Professional, Service, Office, Construction, Production, Drive, Carpool, Transit, Walk, OtherTransp, WorkAtHome, MeanCommute, Employed, PrivateWork, PublicWork, SelfEmployed, FamilyWork, Unemployment
- What do you think should be included as the inputs?


```{r}

```

## Data preparation

Here we need to prepare the data, in particular:

1. Train-test split
2. Data preprocessing

Using `caret` (or something else if you like), prepare two datasets of pre-processed train/test data.

## Train-test split

```{r}

```

## Preprocess

```{r}

```


## Analysis

### Linear regression

- Run linear regression 
- Evaluate the model

```{r}

```
```{r}

```

### Additional movel evaluations

Using the linear regression model as the baseline we attempt two things:

1. Is it possible to improve the prediction using more flexible models?
  - KNN-regression
  - Or regression model variant of models covered in classificaiton section. 
    - For example:
      - svm: svmPoly, svmRadial works both regression and classification (svmPoly may take quite long time as the number of tuning paramters are many.)
      - trees: rf
      


```{r}

```

### SVM with Radial Kernel

```{r}

```


## LASSO and ridge regression

- Now, let's run LASSO and/or Ridge regression. 
- What do you find? 
  - Shrinkage of the coefficients

### LASSO Outcome

```{r}

```

#### Plot with `plot_glmnet`

Shrinkage plot of `glmnet` is not informative as it won't show the variable name. Instead you can use `plot_glmnet` in `plotmo` package.

```{r}

```



### Ridge regression outcome

```{r}

```

#### Plot with `plot_glmnet`

```{r}

```

### Compare coefs: lm, lasso/ridge

Compare the cefficients across the models. What do you find?

```{r}

```